---
title: Testing Effects of Transport Media Reagents
output: github_document
---
## Cytokines measured

IFN-α2a, IL-17A, IL-18, IL-1α, IL-7, IP-10   
MIP-1α, MIP-1β, MIP-3α, IFNg, IL-10, IL-12p70
IL-1β, IL-2, IL-6, IL-8, TNF-α   

```{r read in 7Dec16 data, echo = FALSE, message = FALSE , warning = FALSE}


library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)


#We ran some samples for Josh Schiffer's group on this MSD run so I'll analyze first.


dat7Dec16 <- read.csv("../Sample Runs/7Dec16 9plex_8plex/raw data from DiscoveryWorkbench/7Dec16 8plex_9plex.csv", skip = 1)



#Fix the greek

dat7Dec16$Assay <-recode(dat7Dec16$Assay,
                                      "IFN-Î±2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1Î±" = "IL-1\U03B1",
                                      "MIP-1Î±"="MIP-1\U03B1",
                                      "TNF-Î±" = "TNF-\U03B1",
                                      "MIP-3Î±" = "MIP-3\U03B1",
                                      "MIP-1Î²" = "MIP-1\U03B2",
                                      "IL-1Î²" = "IL-1\U03B2",
                                      "IFN-Î³" = "IFNg")


# forJosh<-dat7Dec16 %>%
#   filter(Sample.Group == "Standards"|str_detect(Sample, "P|pool"))%>%
#   mutate(Dilution = ifelse(str_detect(Sample, "1:10"),"1:10","neat"))%>%
#   write.csv(file = "raw_data_for_Josh.csv")

```

## Checking Standard Recovery and Curves


Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards,echo=FALSE, message = FALSE, warning = FALSE}


stnds_dat7Dec16<-filter(dat7Dec16, Sample.Group == "Standards")



#which standards are out of det. range?
standardsOutOfRange<-stnds_dat7Dec16 %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)



#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery) 

stnds_below_dat7Dec16<- stnds_dat7Dec16%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())


whichAssay_stnds_below_dat7Dec16<-stnds_dat7Dec16 %>%
   filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  filter(!str_detect(Sample, "008"))%>%
  select(Sample, Assay, Detection.Range)
```


```{r percent recovery, fig.height = 7, fig.width = 12,echo=FALSE, message = FALSE, warning = FALSE}
#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

stnds_percent_recov_dat7Dec16<- stnds_dat7Dec16 %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)


#plotting percent recovery with lines showing the bounds of "good recovery" (btwn 80 and 120%)

ggplot(stnds_percent_recov_dat7Dec16,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 180))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = 9))+
  ggtitle("Percent Recovery of Standards")
```



```{r rsquared,echo=FALSE, message = FALSE, warning = FALSE}

#checking the R squared goodness-of-fit for the standard curves
rSquared<-stnds_percent_recov_dat7Dec16%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))
```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from the least concentrated standard (diluent only) except for two from Standard 6 and five from standard 7:

`r pander(stnds_below_dat7Dec16)`

`r pander(whichAssay_stnds_below_dat7Dec16)`

```{r Josh samples and controls,echo=FALSE, message = FALSE, warning = FALSE}

#make a column for Dilution and a column for sample type

CVLpool<-dat7Dec16 %>%
filter(str_detect(Sample, "P|pool"))%>%
mutate(Dilution = ifelse(
str_detect(Sample, "1:10"), "1:10", "neat"))%>%
mutate(Type = ifelse(
str_detect(Sample, "Cocktail"), "Test","Control"))
```


## Testing effects of transport media cocktail

We added the following reagents to our usual control CVL pool to mimic the transport medium:

* Protease Inhibitor at a 1x final concentration (Calbiochem, lot 2746008, catalog #: 539131-1VL)

* 10% Igepal (Sigma, lot 51K0084, catalog #: 1-3021)

* 0.25% BSA (Sigma, lot 018K699, cat # A9647-500G)

We tested both a "neat" and 1:10 dilution of the samples. All samples were run in duplicate.

```{r below detection,echo=FALSE, message = FALSE, warning = FALSE}

CVLpoolDetect<-CVLpool %>%
filter(!Detection.Range == "In Detection Range")%>%
select(Assay, Dilution,Type, Detection.Range)%>%
group_by(Assay,Type,Dilution, Detection.Range)%>%
summarise("Samples" = n())




missingDiluted <- CVLpoolDetect %>%
  filter(Dilution == "1:10", Samples == 2)

missingNeat<-CVLpoolDetect %>%
  filter(Dilution == "neat", Samples == 2)

```
  
#Results

For some cytokines, some of the sample concentrations fell below the detection range:

`r pander(CVLpoolDetect)`


There was a wide range of concentrations across cytokines but the test and control samples were similar. Below are plots showing the calculated concentrations for the *1:10 diluted* and *neat* samples. 

Note that the concentrations are plotted on a *log 10* scale.

Some conditions are missing dots: these are ones where the concentration was below the level of detection.



```{r  plot diluted,echo=FALSE, message = FALSE, warning = FALSE}


inRangeCVLpool<-CVLpool %>%
filter(Detection.Range == "In Detection Range")

#diluted samples
diluted<-inRangeCVLpool %>%
  filter(Dilution == "1:10")

ggplot(diluted, aes(x = Type, y = log10(Calc..Concentration)))+
geom_jitter(aes(), width = 0.25, alpha = 0.5, size = 3)+
labs(y = "log10 Cytokine Concentration (pg/ml)")+
facet_wrap(~Assay)+
ggtitle("log10 Cytokine Concentrations of *1:10 diluted* samples")
```

```{r  plot neat,echo=FALSE, message = FALSE, warning = FALSE}
#neat samples 

neat<-inRangeCVLpool %>%
  filter(Dilution != "1:10")

ggplot(neat, aes(x = Type, y = log10(Calc..Concentration)))+
geom_jitter(aes(), width = 0.25, alpha = 0.5, size = 3)+
labs(y = "log10 Cytokine Concentration (pg/ml)")+
facet_wrap(~Assay)+
ggtitle("log10 Cytokine Concentrations of *neat* samples")
```



```{r plotting standards and samples setup,echo=FALSE, message = FALSE, warning = FALSE}

#appendix C of discovery workbench user guide shows which parameters I need for the FourPL regression that was used for drawing the standard curve.

#y = b1 + (b2-b1)/(1+(x/b3)^b4)
# Where:
# y=ECL counts (response)
# x=Concentration (known calibrator concentration)
# b1=Maximum response plateau (Calc. Top)
# b2=Minimum response plateau (Calc. Bottom)
# b3=Concentration at which 50% of the maximal response is observed (Calc. MidPoint)
# b4="Slope" or shape parameter (Calc. HillSlope)

plotStandards<-dat7Dec16 %>%
  filter(Sample.Group == "Standards") %>% 
  select(Sample,Assay,Signal,Concentration,Algorithm.Parameter..Calc..Top,Algorithm.Parameter..Calc..Bottom,Algorithm.Parameter..Calc..MidPoint,Algorithm.Parameter..Calc..HillSlope)

#make better names
names(plotStandards)<-c("Sample","Assay","Signal","Concentration","CalcTop","CalcBottom","CalcMid","CalcHillslope")

#remove std 8 (discovery workbench doesn't include it and I"m trying to match my plots to them except less ugly geez.)
plotStandardsNoS008<-plotStandards %>% 
  filter(Sample != "S008")
```

```{r plotting standards and samples,echo=FALSE, message = FALSE, warning = FALSE, fig.height= 10, fig.width = 10}
#plotting the curves for the standards using the FourPL equation provided by Discovery Workbench with the samples overlaid on the curve.

#There are 6 rows containing missing values: these are the values that were "Below Fit Curve Range" so DWB only gives me "NaN" for the caluclated concentration.

#There are other samples that were "Below Detection Range", but I still get values for those, just not reliable ones.


ggplot(plotStandardsNoS008,aes(x = log10(Concentration), y =
log10(CalcTop + ((CalcBottom - CalcTop)/(1+ (Concentration/CalcMid)^CalcHillslope)))))+
  geom_line(aes(group = Assay),size = .2, alpha = 0.5, linetype = "dashed")+
  geom_point(data = CVLpool, aes(x = log10(Calc..Concentration), y = log10(Signal), group = Assay, color = Type, shape = Dilution),size = 3, alpha = 0.5)+
  scale_shape_manual(values = c("neat"= 16,"1:10" = 1 ))+
  facet_wrap(~Assay)+
  xlim(c(-2.5, 5.5))+
  labs(y = "log10 (Signal)")+
  ggtitle("Standard Curves and Samples")+
  theme(strip.text = element_text(size =12))
```  
 