---
title: Testing Effects of Transport Media Reagents
output: github_document
---

```{r read in 7Dec16 data, echo = FALSE, message = FALSE , warning = FALSE}


library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)


#We ran some samples for Josh Schiffer's group on this MSD run so I'll analyze first.


dat7Dec16 <- read.csv("../Sample Runs/7Dec16 9plex_8plex/raw data from DiscoveryWorkbench/7Dec16 8plex_9plex.csv", skip = 1)



#Fix the greek

dat7Dec16$Assay <-recode(dat7Dec16$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg")


forJosh<-dat7Dec16 %>%
  filter(Sample.Group == "Standards"|str_detect(Sample, "P|pool"))%>%
  mutate(Dilution = ifelse(str_detect(Sample, "1:10"),"1:10","neat"))%>%
  write.csv(file = "raw_data_for_Josh.csv")

```

## Checking Standard Recovery and Curves


Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards,echo=FALSE, message = FALSE, warning = FALSE}


stnds_dat7Dec16<-filter(dat7Dec16, Sample.Group == "Standards")



#which standards are out of det. range?
standardsOutOfRange<-stnds_dat7Dec16 %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)



#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery) 

stnds_below_dat7Dec16<- stnds_dat7Dec16%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())


whichAssay_stnds_below_dat7Dec16<-stnds_dat7Dec16 %>%
   filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  filter(!str_detect(Sample, "008"))%>%
  select(Sample, Assay, Detection.Range)
```


```{r percent recovery, fig.height = 7, fig.width = 12,echo=FALSE, message = FALSE, warning = FALSE}
#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

stnds_percent_recov_dat7Dec16<- stnds_dat7Dec16 %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)


#plotting percent recovery with lines showing the bounds of "good recovery" (btwn 80 and 120%)

ggplot(stnds_percent_recov_dat7Dec16,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 180))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = 9))+
  ggtitle("Percent Recovery of Standards")
```



```{r rsquared,echo=FALSE, message = FALSE, warning = FALSE}

#checking the R squared goodness-of-fit for the standard curves
rSquared<-stnds_percent_recov_dat7Dec16%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))
```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from the least concentrated standard (diluent only) except for two from Standard 6 and five from standard 7:

`r pander(stnds_below_dat7Dec16)`

`r pander(whichAssay_stnds_below_dat7Dec16)`

```{r Josh samples and controls,echo=FALSE, message = FALSE, warning = FALSE}

#make a column for Dilution and a column for sample type

CVLpool<-dat7Dec16 %>%
filter(str_detect(Sample, "P|pool"))%>%
mutate(Dilution = ifelse(
str_detect(Sample, "1:10"), "1:10", "neat"))%>%
mutate(Type = ifelse(
str_detect(Sample, "Cocktail"), "Test","Control"))
```


## Testing effects of transport media cocktail

We added the following reagents to our usual control CVL pool to mimic the transport medium:

* Protease Inhibitor at a 1x final concentration (Calbiochem, lot 2746008, catalog #: 539131-1VL)

* 10% Igepal (Sigma, lot 51K0084, catalog #: 1-3021)

* 0.25% BSA (Sigma, lot 018K699, cat # A9647-500G)

We tested both a "neat" and 1:10 dilution of the samples. All samples were run in duplicate.

```{r below detection,echo=FALSE, message = FALSE, warning = FALSE}

CVLpoolDetect<-CVLpool %>%
filter(!Detection.Range == "In Detection Range")%>%
select(Assay, Dilution,Type, Detection.Range)%>%
group_by(Assay,Type,Dilution, Detection.Range)%>%
summarise("Number In Range" = n())

names(CVLpoolDetect)[4]<- "Detection Range"

```
  
#Results

For some cytokines, some of the sample concentrations fell below the detection range:

`r pander(CVLpoolDetect)`


There was a wide range of concentrations across cytokines but the test and control samples were similar:

```{r all concentrations,echo=FALSE, message = FALSE, warning = FALSE}


inRangeCVLpool<-CVLpool %>%
filter(Detection.Range == "In Detection Range")


ggplot(inRangeCVLpool, aes(x = Dilution, y = Calc..Concentration))+
geom_point(aes(color = Type))+
labs(y = "Cytokine Concentration (pg/ml)")+
facet_wrap(~Assay, scales = "free_y")+
ggtitle("Cytokine Concentrations (note scales)")

```


  